---
title: "Hands-on Exercise 6"
description: |
  Spatial Weights and Applications
author:
  - name: Yu Yiling
    url: https://www.linkedin.com/in/yiling-yu/
date: 09-20-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Install packages

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# 2. Getting the Data Into R Environment

## a. Import shapefile into r environment
```{r}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")
```

## b. Import csv file into r environment
```{r}
hunan2012 <- read_csv("data/attribute/Hunan_2012.csv")
```

## c. Performing relational join
```{r}
hunan <- left_join(hunan,hunan2012)
```

## d. Visualising Regional Development Indicator
```{r}
qtm(hunan, "GDPPC")
```

# 3. Computing Contiguity Spatial Weights

## a. Computing (QUEEN) contiguity based neighbours/weight matrix
```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

* The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one heighbours.

* For each polygon in our polygon object, wm_q lists all neighboring polygons.

```{r}
# To see the neighbors for the first polygon in the object
wm_q[[1]]
```

* Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

```{r}
#Retrieve the county name of Polygon ID=1
hunan$County[1]

# Reveal the county names of the five neighboring polygons
hunan$NAME_3[c(2,3,4,57,85)]

# display the complete weight matrix
#str(wm_q)
```

## b. Creating (ROOK) contiguity based neighbours/weight matrix
```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

* The summary report above shows that there are 88 area units in Hunan. The most connect area unit has 10 neighbours. There are two area units with only one heighbours. Why different?

## c. Visualising contiguity weights

* A connectivity graph takes a point and displays a line to each neighboring point. We are working with polygons at the moment, so we will need to get points in order to make our connectivity graphs. The most typically method for this will be polygon centroids. We will calculate these in the sf package before moving onto the graphs. Getting Latitude and Longitude of Polygon Centroids

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
head(coords)
```

#### Plotting Queen contiguity based neighbours maps
```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

#### Plotting Rook contiguity based neighbours maps
```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

#### compare both maps
```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

## d. Computing distance based neighbours/weight matrices

* The function identifies neighbours of region points by Euclidean distance with a distance band with lower d1= and upper d2= bounds controlled by the bounds= argument. If unprojected coordinates are used and either specified in the coordinates object x or with x as a two column matrix and longlat=TRUE, great circle distances in km will be calculated assuming the WGS84 reference ellipsoid.

#### Determine the cut-off distance
```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

* The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

#### Computing fixed distance weight matrix
```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62

# display the content of wm_d62 weight matrix
# str(wm_d62)
# table(hunan$County, card(wm_d62))

n_comp <- n.comp.nb(wm_d62)
n_comp$nc
table(n_comp$comp.id)
```

#### Plotting fixed distance weight matrix
```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

* The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

#### Computing adaptive distance weight matrix

* One of the characteristics of fixed distance weight matrix is that more densely settled areas (usually the urban areas) tend to have more neighbours and the less densely settled areas (usually the rural counties) tend to have lesser neighbours. Having many neighbours smoothes the neighbour relationship across more neighbours.
* It is possible to control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6)) #each county to have six neighbours!
knn6
# str(knn6)
```

#### Plotting distance based neighbours/weight matrix
```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## e. Computing distance based neighbours/weight matrices
#### compute the inverse distances
```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
head(ids)
```

#### Row-standardised weights matrix: equal weight
```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

* We need to assign weights to each neighboring polygon. In our case, each neighboring polygon will be assigned equal weight (style=“W”). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. While this is the most intuitive way to summaries the neighbors’ values it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we’ll stick with the style=“W” option for simplicity’s sake but note that other more robust options are available, notably style=“B”.
* The zero.policy=TRUE option allows for lists of non-neighbors. This should be used with caution since the user may not be aware of missing neighbors in their dataset however, a zero.policy of FALSE would return an error.

```{r}
# To see the weight of the first polygon’s four neighbors type
rswm_q$weights[10]
```

* Each neighbor is assigned a 0.2 of the total weight. This means that when R computes the average neighboring income values, each neighbor’s income will be multiplied by 0.2 before being tallied. Is there any typo here?

#### Row-standardised weights matrix: Inverse Distance Weight (IDW)
```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids

rswm_ids$weights[1]
summary(unlist(rswm_ids$weights))
```

# 4. Application of Spatial Weight Matrix

## a. Spatial lag with row-standardized weights
#### average neighbor GDPPC value for each polygon = spatially lagged values
```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

#### append the spatially lag GDPPC values onto hunan SpatialPolygonDataFrame
```{r}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
# head(hunan)
```

#### plot both the GDPPC and spatial lag GDPPC for comparison
```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```

## b. Spatial window sum

* The spatial window sum uses and includes the diagonal element.

#### assign knn6 to a new variable because we will directly alter its structure to add the diagonal elements
```{r}
knn6a <- knn6
```

#### To add the diagonal element to the neighbour list
```{r}
include.self(knn6a)
```

#### assign binary weights to the neighbour structure that includes the diagonal element
```{r}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```

#### explicitly assign weight values
```{r}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")
```

#### compute the lag variable
```{r}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)
```

#### convert the lag variable listw object into a data.frame
```{r}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC") #renames the field names of lag_knn6.res object into NAME_3 and lag_sum GDPPC respectively
```

#### append lag_sum GDPPC values onto hunan sf data.frame
```{r}
hunan <- left_join(hunan, lag_knn6.res)
```

#### plot the GDPPC and lag_sum GDPPC map for comparison
```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```






