---
title: "Hands-on Exercise 2"
description:

author:
  - name: Yu Yiling
    url: https://www.linkedin.com/in/yiling-yu/
date: 09-05-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
  * install packages
  * import data: st_read()
  * peek data: st_geometry(), glimpse(), head()
  * projection transformation: st_crs(), st_set_crs(), st_transform()
  * Import and convert an aspatial data: read_csv(), st_as_sf(), st_transform()
  * sf package: buffering, point-in-polygon count
  * EDA: hist(), ggplot()

# 1. Install packages

```{r}
packages = c('sf', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# 2. Import data

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
                  
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPath")

preschool = st_read("data/geospatial/pre-schools-location-kml.kml")
```

# 3. Peek the data

```{r}
st_geometry(mpsz)
glimpse(mpsz)
head(mpsz, n=5) 
plot(mpsz)
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```

# 4. Projection transformation

```{r}
st_crs(mpsz)
mpsz3414 <- st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
cyclingpath3414 <- st_set_crs(cyclingpath, 3414)
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
st_geometry(preschool3414)
```

# 5. Import and convert an aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
list(listings) 
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
glimpse(listings_sf)
```

# 6. Geoprocessing with sf package

## a. buffering

#### Q: "The authority is planning to upgrade the exiting cycling path. To do so, they need to acquire 5 metres of reserved land on the both sides of the current cycling path. You are tasked to determine the extend of the land need to be acquired and their total area."

#### Solution:
```{r}
buffer_cycling <- st_buffer(cyclingpath3414, 
                               dist=5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

## b. Point-in-polygon count

#### Q: "A pre-school service group want to find out the numbers of pre-schools in each Planning Subzone."

#### Solution:
```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
```

#### Q: "Calculate the density of pre-school by planning subzone."

#### Solution:
```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# 7. EDA

```{r}
hist(mpsz3414$`PreSch Density`)

ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

#### Q: "Using ggplot2 method, plot a scatterplot showing the relationship between Pre-school Density and Pre-school Count."

#### Solution: 
```{r}
#???
```

