---
title: "In-class Exercise 4"
description:
author:
  - name: Yu Yiling
    url: https://www.linkedin.com/in/yiling-yu/
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
  * install packages
  * import data and transform aspatial data to geospatial data: st_read(), read_rds(), st_as_sf()
  * data Wrangling: sf -> Spatial class -> sp -> spatstat ppp format
  * plot map

# 1. Install Packages

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap','tidyverse')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

# 2. Import data and transform aspatial data

```{r}
sg_sf <- st_read("data/shapefile", layer="CostalOutline")
mpsz_sf <- st_read(dsn = "data/shapefile", 
                layer = "MP14_SUBZONE_WEB_PL")

childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```
```{r}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"),
                    crs=3414)
```
```{r}
childcare$Lat <- as.numeric(childcare$Lat)
childcare$Lng <- as.numeric(childcare$Lng)
childcare_sf <- st_as_sf(childcare,
                         coords = c("Lng",
                                    "Lat"),
                         crs=4326) %>%
  st_transform(crs =3414)
```

# 3. Data Wrangling

## a. Converting from sf to Spatial class

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

## b. Converting from Spatial class to sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

## c. Converting from sp to spatstat ppp format

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

# 4. Plot map

```{r}
tmap_mode('view')
tm_shape(childcare_sf) +
  tm_dots(alpha=0.4,
          col="blue",
          size=0.05) +
tm_shape(CHAS_sf) +
  tm_dots(alpha=0.4,
          col="red",
          size=0.05)
```



