---
title: "Hands-on Exercise 11"
description: |
  Modelling Geographical Accessibility
author:
  - name: Yu Yiling
    url: https://www.linkedin.com/in/yiling-yu/
date: 11-01-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Install packages
```{r}
packages = c('tmap', 'SpatialAcc',
             'sf', 'ggstatsplot', 'reshape2',
             'tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# 2. Import and prepare data
## a. geospatial data
```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 

mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)

st_crs(mpsz)
st_crs(eldercare)
st_crs(hexagons)
```

#### Cleaning and updating attribute fields of the geospatial data
```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)

hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

* Notice that for the purpose of this hands-on exercise, a constant value of 100 is used. In practice, actual demand of the hexagon and capacity of the eldercare centre should be used.

## b. aspatial data
```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

#### Tidying distance matrix

* To transform the O-D matrix in to a matrix that the rows represent origins (i.e. also know as from field) and the columns represent destination (i.e. also known as to field.)

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

#### convert the unit f measurement from metre to kilometre
```{r}
distmat_km<-as.matrix(distmat/1000)
```

# 3. Modelling and Visualising Accessibility using Hansen Method
## a. Computing Hansen’s accessibility
```{r}
acc_Hansen <- data.frame(ac(hexcen$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))
colnames(acc_Hansen) <- "accHansen" #rename
acc_Hansen <- tbl_df(acc_Hansen) #convert the data table into tibble format
```

#### join the acc_Hansen tibble data frame with the hexagons simple feature data frame
```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

## b. Visualising Hansen’s accessibility
#### Extracting map extend
```{r}
mapex <- st_bbox(hexagons) #view extent
```

#### create a high cartographic quality accessibility to eldercare centre in Singapore
```{r}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## c. Statistical graphic visualisation

* In this section, we are going to compare the distribution of Hansen’s accessibility values by URA Planning Region.

#### add the planning region field into haxegon_Hansen simple feature data frame
```{r}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

#### plot the distribution by using boxplot graphical method
```{r}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# 4. Modelling and Visualising Accessibility using KD2SFCA Method
## a. Computing KD2SFCA’s accessibility
```{r}
acc_KD2SFCA <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "KD2SFCA"))

colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- tbl_df(acc_KD2SFCA)
```

#### join the acc_KD2SFCA tibble data frame with the hexagons simple feature data frame
```{r}
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## b. Visualising KD2SFCA’s accessibility
#### create a high cartographic quality accessibility to eldercare centre in Singapore
```{r}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## c. Statistical graphic visualisation

* Now, we are going to compare the distribution of KD2CFA accessibility values by URA Planning Region.

#### add the planning region field into hexagon_KD2SFCA simple feature data frame
```{r}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```

#### plot the distribution by using boxplot graphical method
```{r}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# 5. Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method
## a. Computing SAM accessibility
```{r}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))

colnames(acc_SAM) <- "accSAM"
acc_SAM <- tbl_df(acc_SAM)
```

#### join the acc_SAM tibble data frame with the hexagons simple feature data frame
```{r}
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## b. Visualising SAM’s accessibility
#### create a high cartographic quality accessibility to eldercare centre in Singapore
```{r}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## c. Statistical graphic visualisation

* Now, we are going to compare the distribution of KD2CFA accessibility values by URA Planning Region.

#### add the planning region field into hexagon_SAM simple feature data frame
```{r}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

#### plot the distribution by using boxplot graphical method
```{r}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```




