---
title: "In-class Exercise 5"
description:
author:
  - name: Yu Yiling
    url: https://www.linkedin.com/in/yiling-yu/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3)
```

# 1. Install packages

```{r}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse','plotly', 'ggthemes')
for(p in packages){
  if(!require(p, character.only=T)){
    install.packages(p)
  }
library(p,character.only = T)
}
```

* Note:
  + Maptool is for creating spatical objects
  + sf is for importing geospatial data
  + raster is for creating raster object
  + spatstat is for performing SPPA
  + tmap is for plotting thematic maps
  + tidyverse is for data handling 

# 2. Import data

```{r}
CHAS <- read_rds("data/rds/CHAS.rds")
childcare <- read_rds("data/rds/childcare.rds")

mpsz_sf <- st_read(dsn = "data/shapefile", layer = "MP14_SUBZONE_WEB_PL")
```
```{r}
glimpse(CHAS)
glimpse(childcare)

mpsz_sf <- st_set_crs(mpsz_sf, 3414)
st_crs(mpsz_sf)
```

* WGS84 for childcare dataset is assumed. why?

# 3. Coverting the aspatial data frame into sf objects

```{r}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"),
                    crs=3414)
```

* Note: *st_as_sf()* accept coordinate in character data type

```{r}
childcare_sf <- st_as_sf(childcare,
                    coords = c("Lng",
                               "Lat"),
                    crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
#plot the graph to have a look
tmap_mode("view")
tm_shape(childcare_sf) +
  tm_dots(alpha=0.4,
          col= "blue",
          size = 0.05) +
tm_shape(CHAS_sf) +
  tm_dots(alpha=0.4,
          col= "red",
          size = 0.05)
```

# 4. Convert sf to ppp objects

## a. Convert sf data frame to Spatial class
```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

## b. Convert Spatial class into generic sp object
```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

* as.SpatialPoint() of as.SpatialPolygon() of maptools package 

## c. Covert generic sp object into ppp object
```{r}
childcare_ppp <- as(childcare_sp,"ppp")
CHAS_ppp <- as(CHAS_sp,"ppp")
```

* using as.ppp() of maptools package.

## d. Remove duplicate points using jitter
```{r}
childcare_ppp_jit <- rjitter(childcare_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)
any(duplicated(childcare_ppp_jit))

CHAS_ppp_jit <- rjitter(CHAS_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)
any(duplicated(CHAS_ppp_jit))
```

# 5. Extracting Punggol Planning Area

```{r}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

# 6. Creating owin object

## a. Convert SpatialPolygonsDataFrame into SpatialPolygons object
```{r}
pg_sp <- as(pg, "SpatialPolygons")
```

## b. Convert SpatialPolygons into owin object
```{r}
pg_owin <- as(pg_sp, "owin")
```

## c. Extract spatial points window owin
```{r}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]

plot(childcare_ppp_jit)
plot(childcare_pg)
```

# 7. Analysing Spatial Point Process Using L-Function

## Monte Carlo test with L-fucntion
```{r}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim = 99,
                        rank = 1,
                        gloval = TRUE)

L_CHAS <- envelope(CHAS_pg,
                   Lest,
                   nsim = 99,
                   rank = 1,
                   gloval = TRUE)
```

## plotting interactive L-function

#### L_childcare
```{r eval=FALSE}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```

#### L_CHAS
```{r eval=FALSE}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_CHAS)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
}
```


